### 2016-03-12 FontTest/fonttest.tex

\chapter*{Font test}

\emph{textrm:}
\textrm{On the other hand, we denounce with righteous indignation and
  dislike men who are so beguiled and demoralized by the charms of
  pleasure of the moment, so blinded by desire, that they cannot
  foresee the pain and trouble that are bound to ensue; and equal
  blame belongs to those who fail in their duty through weakness of
  will, which is the same as saying through shrinking from toil and
  pain. These cases are perfectly simple and easy to distinguish. In a
  free hour, when our power of choice is untrammelled and when nothing
  prevents our being able to do what we like best, every pleasure is
  to be welcomed and every pain avoided. But in certain circumstances
  and owing to the claims of duty or the obligations of business it
  will frequently occur that pleasures have to be repudiated and
  annoyances accepted. The wise man therefore always holds in these
  matters to this principle of selection: he rejects pleasures to
  secure other greater pleasures, or else he endures pains to avoid
  worse pains.}

\pagebreak

\input{FontTest/textfragment}

\pagebreak 
\emph{textsf:}
\textsf{On the other hand, we denounce with righteous indignation and
  dislike men who are so beguiled and demoralized by the charms of
  pleasure of the moment, so blinded by desire, that they cannot
  foresee the pain and trouble that are bound to ensue; and equal
  blame belongs to those who fail in their duty through weakness of
  will, which is the same as saying through shrinking from toil and
  pain. These cases are perfectly simple and easy to distinguish. In a
  free hour, when our power of choice is untrammelled and when nothing
  prevents our being able to do what we like best, every pleasure is
  to be welcomed and every pain avoided. But in certain circumstances
  and owing to the claims of duty or the obligations of business it
  will frequently occur that pleasures have to be repudiated and
  annoyances accepted. The wise man therefore always holds in these
  matters to this principle of selection: he rejects pleasures to
  secure other greater pleasures, or else he endures pains to avoid
  worse pains.}

\emph{texttt:}
\texttt{On the other hand, we denounce with righteous indignation and
  dislike men who are so beguiled and demoralized by the charms of
  pleasure of the moment, so blinded by desire, that they cannot
  foresee the pain and trouble that are bound to ensue; and equal
  blame belongs to those who fail in their duty through weakness of
  will, which is the same as saying through shrinking from toil and
  pain. These cases are perfectly simple and easy to distinguish. In a
  free hour, when our power of choice is untrammelled and when nothing
  prevents our being able to do what we like best, every pleasure is
  to be welcomed and every pain avoided. But in certain circumstances
  and owing to the claims of duty or the obligations of business it
  will frequently occur that pleasures have to be repudiated and
  annoyances accepted. The wise man therefore always holds in these
  matters to this principle of selection: he rejects pleasures to
  secure other greater pleasures, or else he endures pains to avoid
  worse pains.}

\emph{textit:}
\textit{On the other hand, we denounce with righteous indignation and
  dislike men who are so beguiled and demoralized by the charms of
  pleasure of the moment, so blinded by desire, that they cannot
  foresee the pain and trouble that are bound to ensue; and equal
  blame belongs to those who fail in their duty through weakness of
  will, which is the same as saying through shrinking from toil and
  pain. These cases are perfectly simple and easy to distinguish. In a
  free hour, when our power of choice is untrammelled and when nothing
  prevents our being able to do what we like best, every pleasure is
  to be welcomed and every pain avoided. But in certain circumstances
  and owing to the claims of duty or the obligations of business it
  will frequently occur that pleasures have to be repudiated and
  annoyances accepted. The wise man therefore always holds in these
  matters to this principle of selection: he rejects pleasures to
  secure other greater pleasures, or else he endures pains to avoid
  worse pains.}

\emph{textsf:}
\textsl{On the other hand, we denounce with righteous indignation and
  dislike men who are so beguiled and demoralized by the charms of
  pleasure of the moment, so blinded by desire, that they cannot
  foresee the pain and trouble that are bound to ensue; and equal
  blame belongs to those who fail in their duty through weakness of
  will, which is the same as saying through shrinking from toil and
  pain. These cases are perfectly simple and easy to distinguish. In a
  free hour, when our power of choice is untrammelled and when nothing
  prevents our being able to do what we like best, every pleasure is
  to be welcomed and every pain avoided. But in certain circumstances
  and owing to the claims of duty or the obligations of business it
  will frequently occur that pleasures have to be repudiated and
  annoyances accepted. The wise man therefore always holds in these
  matters to this principle of selection: he rejects pleasures to
  secure other greater pleasures, or else he endures pains to avoid
  worse pains.}

\emph{textsc:}
\textsc{On the other hand, we denounce with righteous indignation and
  dislike men who are so beguiled and demoralized by the charms of
  pleasure of the moment, so blinded by desire, that they cannot
  foresee the pain and trouble that are bound to ensue; and equal
  blame belongs to those who fail in their duty through weakness of
  will, which is the same as saying through shrinking from toil and
  pain. These cases are perfectly simple and easy to distinguish. In a
  free hour, when our power of choice is untrammelled and when nothing
  prevents our being able to do what we like best, every pleasure is
  to be welcomed and every pain avoided. But in certain circumstances
  and owing to the claims of duty or the obligations of business it
  will frequently occur that pleasures have to be repudiated and
  annoyances accepted. The wise man therefore always holds in these
  matters to this principle of selection: he rejects pleasures to
  secure other greater pleasures, or else he endures pains to avoid
  worse pains.}

\emph{textbf:}
\textbf{On the other hand, we denounce with righteous indignation and
  dislike men who are so beguiled and demoralized by the charms of
  pleasure of the moment, so blinded by desire, that they cannot
  foresee the pain and trouble that are bound to ensue; and equal
  blame belongs to those who fail in their duty through weakness of
  will, which is the same as saying through shrinking from toil and
  pain. These cases are perfectly simple and easy to distinguish. In a
  free hour, when our power of choice is untrammelled and when nothing
  prevents our being able to do what we like best, every pleasure is
  to be welcomed and every pain avoided. But in certain circumstances
  and owing to the claims of duty or the obligations of business it
  will frequently occur that pleasures have to be repudiated and
  annoyances accepted. The wise man therefore always holds in these
  matters to this principle of selection: he rejects pleasures to
  secure other greater pleasures, or else he endures pains to avoid
  worse pains.}

\emph{textmd:}
\textmd{On the other hand, we denounce with righteous indignation and
  dislike men who are so beguiled and demoralized by the charms of
  pleasure of the moment, so blinded by desire, that they cannot
  foresee the pain and trouble that are bound to ensue; and equal
  blame belongs to those who fail in their duty through weakness of
  will, which is the same as saying through shrinking from toil and
  pain. These cases are perfectly simple and easy to distinguish. In a
  free hour, when our power of choice is untrammelled and when nothing
  prevents our being able to do what we like best, every pleasure is
  to be welcomed and every pain avoided. But in certain circumstances
  and owing to the claims of duty or the obligations of business it
  will frequently occur that pleasures have to be repudiated and
  annoyances accepted. The wise man therefore always holds in these
  matters to this principle of selection: he rejects pleasures to
  secure other greater pleasures, or else he endures pains to avoid
  worse pains.}

\uppercase{On the other hand, we denounce with righteous indignation
  and dislike men who are so beguiled and demoralized by the charms of
  pleasure of the moment, so blinded by desire, that they cannot
  foresee the pain and trouble that are bound to ensue; and equal
  blame belongs to those who fail in their duty through weakness of
  will, which is the same as saying through shrinking from toil and
  pain. These cases are perfectly simple and easy to distinguish. In a
  free hour, when our power of choice is untrammelled and when nothing
  prevents our being able to do what we like best, every pleasure is
  to be welcomed and every pain avoided. But in certain circumstances
  and owing to the claims of duty or the obligations of business it
  will frequently occur that pleasures have to be repudiated and
  annoyances accepted. The wise man therefore always holds in these
  matters to this principle of selection: he rejects pleasures to
  secure other greater pleasures, or else he endures pains to avoid
  worse pains.}