### 2016-03-12 FrontBackmatter/Acknowledgments.tex

%*******************************************************
% Acknowledgments
%*******************************************************
\begingroup
\let\clearpage\relax
\let\cleardoublepage\relax
\let\cleardoublepage\relax
\chapter*{Acknowledgments}
I would like to express my sincere gratitude to Vas Lubchenko, whose
support often extended well beyond of what is normally required from
the graduate advisor. I am also thankful to him for very careful,
critical, and multiple reading of this manuscript which greatly
contributed to its quality.  This work would not be possible without
help of my family: my parents, my wife, and my son.  I am
deeply grateful to Eric R. Bittner who, by being such a wonderful
graduate advisor to my wife, made an inconspicuous, but very important,
influence on everything I did. I also would like to thank to Anne Goj,
Jon Golden, Pyotr Rabochiy, and other members of Bittner and
Lubchenko's groups who made this place such a nice professional
home. I am indebted to Elena and Phil Motuzko for helping me in the
time of need. Last but not least, I thank Sergei Lukyanets without
whom I would never became a scientist.

The numerical part of this work was completed using TLC2 HPC
facility of the University of Houston and Ares cluster of the Texas
Southern University. I am especially indebted to C. J. Tymczak for
providing the access to the latter.

I also gratefully acknowledge support by the Donors of the ACS Petroleum
Research Fund, the Arnold and Mabel Beckman Foundation, the Welch
Foundation Grant E-1765, and the NSF Grant CHE 0956127.
\endgroup