### 2016-03-29 Chapters/chap1/introduction.tex
Change "Fig. ~" to "Fig.~"

%!TEX root = /Users/tcya/Dropbox/PhDthesis/thesis.tex